\myTitle

### 2023-12-18 GSI_and_Fst.Rmd
Further updates

In order to do this we need to convert the data to a lumped format.
This means making integers of the alleles and then catenating them.
No locus has more than 9 alleles, so we can do single-digits for alleles.  We drop the
LFAR from this, because those were chosen specifically to be weird in the late fall.
```{r}
hfs_ref_long <- full_base2 %>% 
  select(-starts_with("NC_"), -starts_with("Ots")) %>%  # see what we get when we kep just the tag_id markers
  mutate(
    Pop = as.integer(factor(collection, levels = collection_order)),
    .before = collection
  ) %>%
  select(-repunit, -collection, -sample_type) %>%
  pivot_longer(
    cols = c(-indiv, -Pop),
    names_to = c("locus", "gene_copy"),
    names_pattern = "^(.*)_([12])$"
  ) %>%
  group_by(locus) %>%
  mutate(
    alle_int =  as.integer(factor(value)),
    alle_int = ifelse(is.na(alle_int), 0L, alle_int)
  ) %>%
  ungroup() %>%
  select(-value)

# then we can lump the alleles 
hfs_ref_wide <- hfs_ref_long %>%
  group_by(indiv, Pop, locus) %>%
  summarise(geno = str_c(alle_int[1], alle_int[2])) %>%
  ungroup() %>%
  mutate(geno = ifelse(geno == "00", NA_character_, geno)) %>%
  mutate(geno = as.integer(geno)) %>%
  pivot_wider(
    names_from = c(locus),
    values_from = geno
  ) %>%
  arrange(Pop, indiv)

@@

@@

Something is very, very weird here.  Look at that Fst matrix:
```{r}
tmp <- fst_mat
rownames(tmp) <- collection_order
tmp
```

For some reason, BCF, BCS, MDS, and BCF have quite low Fst values to EVERYONE, even Winter run and the
Klamath.  That just cannot be right!!  They have super low Fst values to FRHF and FRHF, but those
two pops have Fst of around 0.14 to winter-run, and BCS to winter run is only 0.0269.  WTF?!

>>>>

In order to do this we need to convert the data to a lumped format.
This means making integers of the alleles and then catenating them.
One locus has 11 allele, so we should use two digits for each allele.  We drop the
LFAR from this, because those were chosen specifically to be weird in the late fall.
```{r}
full_base_long <- full_base2 %>% 
  select(-starts_with("NC_037130.1")) %>%  # drop the LFARs
  mutate(
    Pop = as.integer(factor(collection, levels = collection_order)),
    .before = collection
  ) %>%
  select(-repunit, -collection, -sample_type) %>%
  pivot_longer(
    cols = c(-indiv, -Pop),
    names_to = c("locus", "gene_copy"),
    names_pattern = "^(.*)_([12])$"
  )

hfs_ref_long <- full_base_long %>%
  group_by(locus) %>%
  mutate(
    alle_int =  sprintf("%02d", as.integer(factor(value)))
  ) %>%
  ungroup() %>%
  select(-value)

# then we can lump the alleles 
hfs_ref_wide <- hfs_ref_long %>%
  group_by(indiv, Pop, locus) %>%
  summarise(geno = str_c(alle_int[1], alle_int[2])) %>%
  ungroup() %>%
  mutate(geno = ifelse(str_detect(geno, "NA"), NA_character_, geno)) %>%
  mutate(geno = as.integer(geno)) %>%
  pivot_wider(
    names_from = c(locus),
    values_from = geno
  ) %>%
  arrange(Pop, indiv)

@@

# let's also bootstrap that:
FST_boot <- boot.ppfst(
  dat = hfstat_dat,
  nboot = 1000,
  quant = c(0.025,0.975),
  diploid = TRUE
)

@@

### 2023-12-18 tex/body-text-of-paper.tex
Further updates

Genetic markers have long been used in the 
management of Pacific salmon, and salmon conservation and management has been
a driving force behind a number of advances in molecular ecology,
from data generation techniques \citep{clemento2011discovery,campbell2015genotyping,mckinney2017managing}
to statistical methodology
\citep{smouse1990genetic,anderson2002model,pella2006gibbs}.
One broadly applicable innovation that has been
actively fostered by the Pacific salmon research community is genetic
stock identification
(GSI:~\citealt{milner1982genetic,beacham2004dna,seeb2007development})

@@

@@

 and parentage based tagging
(PBT:~\citealt{anderson2006power, garza2007large, abadia2013large, steele2013validation}).

>>>>

Genetic markers have long been used in the 
management of Pacific salmon, and salmon conservation and management has been
a driving force behind a number of advances in molecular ecology,
from data generation techniques \citep{clemento2011discovery,campbell2015genotyping,mckinney2017managing}
to statistical methodology
\citep{smouse1990genetic,anderson2002model,pella2006gibbs}.
Two broadly applicable innovations that has been
actively fostered by the Pacific salmon research community are genetic
stock identification
(GSI:~\citealt{milner1982genetic,beacham2004dna,seeb2007development})
and parentage based tagging
(PBT:~\citealt{anderson2006power, garza2007large, abadia2013large, steele2013validation}).

@@

A further consideration in the development of a reference baseline for GSI is that
the same panel of loci used for GSI, if properly designed, can also serve as a
markers for parentage based tagging (PBT).

@@
